---
title: "SC1_Proj"
author: "Alessio"
date: "13 January 2020"
output: pdf_document
---

```{r, include=FALSE}
library(ggplot2)
library(magrittr)
library(dplyr)
library(ggcorrplot)
library(tidyr)
library(readr)
library(devtools)
library(tsne)
```

#Description of the dataset and problem
TODO: describe

```{r}
data <- read_csv("../data/data.csv")
```

#Dataset Preprocessing Visualisation and Exploration

```{r}
colSums(is.na(data))
```


```{r}
data %<>% mutate_at(vars(diagnosis), factor)
```

```{r}
train <- data %>% sample_frac(0.8)
test <- anti_join(data,train, by='id')
```

```{r}
data %<>%  
  dplyr::select(-c(id, X33))
train %<>%  
  dplyr::select(-c(id, X33))
test %<>%  
  dplyr::select(-c(id, X33))
```

```{r}
training_data <- train[2:dim(train)[2]] 
training_classes <- train[1] 

test_data <- test[2:dim(test)[2]] 
test_classes <- test[1] 
```


```{r}
sum(is.na(data))
```

```{r}
corr <- data[,-1] %>% 
          cor() %>% 
           round(1)
ggcorrplot(corr,
           hc.order = TRUE,
           colors = c("#6D9EC1", "white", "#E46726"),
           ggtheme = ggplot2::theme_minimal)
```

  - Class Frequencies
  - Density
  - box plots




#Dimensionality Reduction and Feature Selection
  - PCA
Code
```{r}
normalise_z <-  function(X){
  mean_cols <- colMeans(X)
  sd_cols <- apply(X, 2, sd)
  mean_normalised_X <- t(apply(X, 1, function(x){x - mean_cols}))
  normalised_X <- t(apply(mean_normalised_X, 1, function(x){x / sd_cols}))
  return(normalised_X)
}


pca <- function(X, number_components_keep) {
  normalised_X <- normalise_z(X)
  
  corr_mat <- t(normalised_X) %*% normalised_X
  
  eigenvectors <- eigen(corr_mat, symmetric=TRUE)$vectors
  
  reduced_data <-  X %*% eigenvectors[,1:number_components_keep]
  relevant_eigs <-  eigenvectors[,1:number_components_keep]
  returnds <- list(reduced_data, relevant_eigs)
  names(returnds) <- c("reduced_data", "reduction_matrix")
  return(returnds)
}
``` 

Apply to dataset
```{r}
pca_result <- pca(as.matrix(training_data), 2)
reduced_training_data <- data.frame(cbind(pca_result$reduced_data, training_classes))

ggplot(data=reduced_training_data, aes(x=X1, y=X2)) + geom_point(aes(colour=diagnosis))
```
  
 tSNE 
TODO: try different perplexity parameters
```{r}
reduced_training_data <- tsne::tsne(training_data)

reduced_training_data <- data.frame(cbind(reduced_training_data, training_classes))
ggplot(data=reduced_training_data, aes(x=X1, y=X2)) + geom_point(aes(colour=diagnosis))
```

  - Correlation Feature Selection
  - LDA

# Classification
  - Naive Bayes
  - SVM
  - Logistic Regression
  - Lasso

#Conclusion
  - Evaluation of results
  - Discuss outliers