---
title: "Modelling"
author: "Mauro Camara Escudero"
date: "11/14/2019"
output: html_document
---
<!--- For HTML Only --->
`r if (!knitr:::is_latex_output()) '
$\\DeclareMathOperator*{\\argmin}{arg\\,min}$
$\\DeclareMathOperator*{\\argmax}{arg\\,max}$
$\\newcommand{\\var}{\\mathrm{Var}}$
'`
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{amsthm}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{listings}
\usepackage{color}
\usepackage[T1]{fontenc}
\usepackage{fullpage}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{fullpage}
\usepackage{color}
\usepackage{bm}
\usepackage{mathtools}
\usepackage{algorithm2e}
\newcommand{\def}{\overset{\text{def}}{:=}}
\newcommand{\lop}{\mathcal{L}}
\newcommand{\vect}[1]{\boldsymbol{\mathbf{#1}}}
\newcommand{\norm}[1]{||#1||}
\newcommand{\vx}{\vect{x}}
\newcommand{\vb}{\vect{b}}
\newcommand{\vy}{\vect{y}}
\newcommand{\vz}{\vect{z}}
\newcommand{\ve}{\vect{e}}
\newcommand{\yhat}{\widehat{y}}
\newcommand{\xhat}{\widehat{x}}
\newcommand{\vc}{\vect{c}}
\newcommand{\vr}{\vect{r}}
\newcommand{\vphi}{\vect{\phi}}
\newcommand{\vf}{\vect{f}}
\newcommand{\vY}{\vect{Y}}
\newcommand{\vX}{\vect{X}}
\newcommand{\vw}{\vect{w}}
\newcommand{\thetahatis}{\widehat{\theta}^{(s)}_i}
\newcommand{\thetahat}[1]{\widehat{\theta}^{(#1)}_i}
\newcommand{\vm}{\vect{m}}
\newcommand{\redmath}[1]{\mathbin{\textcolor{red}{\vect{#1}}}}
\newcommand{\redtext}[1]{\textcolor{red}{\vect{#1}}}
\newcommand{\vzero}{\vect{0}}
\newcommand{\vt}{\vect{t}}
\newcommand{\linearpredictor}{\vx_i^T\vbeta}
\newcommand{\vmu}{\vect{\mu}}
\newcommand{\vnu}{\vect{\nu}}
\newcommand{\Var}{\text{Var}}
\newcommand{\veta}{\vect{\eta}}
\newcommand{\vbeta}{\vect{\beta}}
\newcommand{\vepsilon}{\vect{\epsilon}}
\newcommand{\Pbb}{\mathbb{P}}
\newcommand{\Ebb}{\mathbb{E}}
\newcommand{\vdelta}{\vect{\delta}}
\newcommand{\Rbb}{\mathbb{R}}
\newcommand{\vxi}{\vect{\xi}}
\newcommand{\vu}{\vect{u}}
\newcommand{\vW}{\vect{W}}
\newcommand{\vlambda}{\vect{\lambda}}
\newcommand{\dataset}{\mathcal{D}}
\newcommand{\vS}{\vect{S}}
\newcommand{\sample}{\vz^{(l)}}
\newcommand{\MYhref}[3][blue]{\href{#2}{\color{#1}{#3}}}
\newcommand{\sol}[1]{\vx^{(#1)}}
\newcommand{\qtext}[1]{\quad\quad \text{#1}}
\newcommand{\vtheta}{\vect{\theta}}
\newcommand{\bi}[1]{\textbf{\textit{#1}}}
\newcommand{\iid}{\overset{\text{i.i.d.}}{\sim}}
\newcommand{\uniform}{\mathcal{U}(0, 1)}
\newcommand{\qimplies}{\quad\Longrightarrow\quad}
\newcommand{\tp}{\tilde{p}}
\newcommand{\nul}{\Theta^{(0)}}
\newcommand{\alter}{\Theta^{(1)}}
\newcommand{\const}{\mathcal{Z}}
\newcommand{\tq}{\tilde{q}}
\newcommand{\vxhat}{\widehat{\vx}}
\newcommand{\tvx}{\widetilde{\vx}}
\newcommand{\tr}{\tilde{r}}
\newcommand{\like}{\mathcal{L}}
\newcommand{\kl}[2]{\text{KL}(#1\,\,||\,\,#2)}
\newcommand{\logit}[1]{\log\left(\frac{#1}{1-#1}\right)}
\newcommand{\elbo}[1]{\text{elbo}(#1)}
\newcommand{\eval}{\biggr\rvert}
\newcommand{\normal}{\mathcal{N}}
\newcommand{\class}{\mathcal{C}}
\newcommand{\infor}{\mathcal{I}}
\newcommand{\variance}{\text{Var}}
\newcommand{\vSigma}{\vect{\Sigma}}
\newcommand{\lp}[3]{
    \begin{equation*}
    \begin{alignat}{2}
    &\!\min       &\qquad& #1\\
    &\text{s.t.} &      & #2\\
    &                  &      & #3
    \end{alignat}
    \end{equation*}
}
\newcommand{\lpmax}[3]{
    \begin{equation*}
    \begin{alignat}{2}
    &\!\max       &\qquad& #1\\
    &\text{s.t.} &      & #2\\
    &                  &      & #3
    \end{alignat}
    \end{equation*}
}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Stochastic Processes
Let $(\Omega, \mathcal{F}, \mathbb{P})$  probability space.

* _Random Variable_: $X:\Omega\to\mathbb{R}$ assigns a number $X(\omega)$ to every outcome $\omega$.
* _Stochastic Process_: $X:T\times\Omega\to\mathbb{R}$ assigns a function $X(t, \omega)$ to every outcome $\omega$.

Basically a stochastic process is a collection of time-dependent functions, that depend on outcomes $\omega$. The following interpretation is useful

* If we fix an outcome $\omega$, then $\vx(t, \omega)$ is a single function of time. Basically a _sample_ of the stochastic process.
* If we fix a time $t$, then $\vx(t, \omega)$ is a _random variable_ representing the _state_ of the process at time $t$.
* If we fix both $\omega$ and $t$ then $\vx(t, \omega)$ is just a number.


```{r echo=FALSE}
library(ggplot2)
library(latex2exp)
library(reshape2)
# set random seed
set.seed(1)
# Define the stochastic process as a function of time and outcome
sp <- function(t, omega){
  return(omega*t + (omega^3)*sin(t) +  omega*cos(1/omega - t) + omega^3)
}
# set a number of total time points
n = 100
# get fluid and static times
t <- seq(0, 10, length.out=n)
t1_val = 5
t2_val = 8
t1 <- rep(t1_val, n/2)
t2 <- rep(t2_val, n/2)
# Set different omegas
o1 = 2.0
o2 = 1.0
o3 = -1.5
# calculate values of function for each static omega, given fluid time
sp1 <- sp(t, o1)
sp2 <- sp(t, o2)
sp3 <- sp(t, o3)
# generate values of function at fixed times t1 and t2
omega <- rnorm(n)  # generate random omega values
spt1 <- sp(t1_val, head(omega, n/2))
spt2 <- sp(t2_val, tail(omega, n/2))
# create dataframe with fluid t and static omegas
df <- data.frame(t=t, omega1=sp1, omega2=sp2, omega3=sp3)
# melt this dataframe
df <- melt(df, id.vars="t")
# create data for fixed time, but fluid omega
variable = c(rep("t1", n/2), rep("t2", n/2))
df2 <- data.frame(t = c(t1,t2), variable=variable, value=c(spt1, spt2))
# put together the dataframes
df_final <- rbind(df, df2)
ggplot(data=df_final, aes(x=t, y=value, color=variable)) + 
  geom_point(data=df2, aes(x=t, y=value, color=variable)) + 
  geom_rug(data=subset(df2, t==t1_val), sides="l", alpha=0.5) + 
  geom_rug(data=subset(df2, t==t2_val), sides="r", alpha=0.5) + 
  geom_line(data=df, aes(x=t, y=value, color=variable)) +
  labs(color="Kept Fixed", x="Time (t)", y=TeX("$x(t, \\omega)$")) + 
  scale_color_discrete(labels=c(expression(omega==2), expression(omega==1), expression(omega==-1.5), expression(t==5),expression(t==8))) + 
  ggtitle("Interpretation of Stochastic Processes") + 
  theme(plot.title = element_text(hjust = 0.5, size=18), axis.title = element_text(size=15)) + 
  guides(color=guide_legend(override.aes = list(linetype=c("solid", "solid", "solid", rep("blank", 2)))))
```

## Non-Probabilistic Classifiers
* SMV to model tectonic boundaries?




