---
title: "Exploratory Analysis"
author: "Andrea Becsek"
date: "7 November 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/home/work/Documents/Project/stats-computing-1")
```

Load libraries
```{r message=FALSE, warning=FALSE}
library(dplyr)
library(readr)
library(ggplot2)
library(maps)
library(reshape2)
library(tidyr)
library(purrr)
library(lubridate)
library(maptools)
library(plotly)
```

Load data
```{r}
df <- read_csv('./data/database.csv')

df_filtered <- df %>%
  filter(!is.na(Time))
df_filtered <- as.data.frame(df_filtered)
```

Replace column names
```{r}
colnames(df_filtered) <- c('date', 'time', 'lat', 'lon', 'type', 'depth', 'depth_er', 'depth_ss', 'magnit', 'magnit_type', 'magnit_er', 'magnit_ss', 'azim', 'horiz_dist', 'horiz_er', 'rms', 'id', 'source', 'loc_source', 'magnit_source', 'status')
```

Convert date to year/month/day format
```{r}
df_filtered <- df_filtered %>% 
  mutate(date = as_date(parse_date_time(date, orders=c('mdy'))))

# add columns for year, month and day
df_filtered <- df_filtered %>% 
  mutate(year = year(date), month = factor(month(date)), day = factor(day(date)))
```


## Dealing with missing values
Percentage of missing values in each column
```{r}
round(100*colMeans(is.na(df_filtered)),2)
```

Drop columns
```{r}
df_filtered <-df_filtered %>% 
  select(-c(depth_er,magnit_er,azim,horiz_er,
            horiz_dist,magnit_type,magnit_ss,depth_ss,id))
```

Replace RMS missing values with the mean
```{r}
df_filtered <- df_filtered %>% 
  mutate(rms=replace(rms, is.na(rms), mean(rms)))
```

Missing value count
```{r}
colSums(is.na(df_filtered))
```


Count number of unique values in each column
```{r}
rapply(df_filtered, function(x) length(unique(x)))
```

## Dealing with categorical variables
```{r}
unique(df_filtered$type)
unique(df_filtered$status)
unique(df_filtered$magnit_source)
unique(df_filtered$loc_source)
unique(df_filtered$source)
```

Store categorical variables as factors
```{r}
categ_var <- c('type', 'status', 'magnit_source', 'loc_source', 'source')
df_filtered[,categ_var] <- lapply(df_filtered[,categ_var],factor)
```

Count observations for each type
```{r}
df_filtered %>% 
  group_by(type) %>% 
  summarise(no_rows = length(type))
```

There are very few explosions and rock bursts so we can remove those observations.

```{r}
df_filtered <- df_filtered %>% 
  filter(grepl('Earthquake|Nuclear', type))
```


We want the variables to contribute equally to the analysis, so we need to make sure that they are on the same scale. So we normalize magnitude, depth and rms

```{r}
normalize <- function(x){
    return((x - min(x)) / (max(x) - min(x)))
}
df_norm <- df_filtered
df_norm[c('magnit','depth','rms')] <- lapply(df_filtered[c('magnit','depth','rms')], normalize)
```

Magnitude classes
```{r}
df_filtered <- df_filtered %>% 
  mutate(class = factor(cut(
     df_filtered$magnit,
     breaks = c(5, 6, 7, 8, Inf),
     labels = c("moderate", "strong", "major", "great"),
     right  = FALSE)
     )
)
```

## Some plots
Histograms for numeric variables
```{r}
df_filtered %>% 
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value))+
  facet_wrap(~key, scales='free')+
  geom_histogram()
```

Histogram of earthquakes
```{r}
df_filtered %>% 
  ggplot(aes(x=magnit))+
  geom_histogram(binwidth=0.5)
```

Count earthquakes for each bin
```{r}
df_filtered %>% 
  count(cut_width(magnit,0.5))
```


Number of earthquakes for each magnitude class
```{r}
df_filtered %>% 
  group_by(class) %>% 
  summarise(events = n()) %>% 
  ggplot(aes(x = class, y = events))+
  geom_bar(stat = 'identity')
```

Lets zoom in and focus only on moderate and strong earthquakes
```{r}
df_filtered %>% 
  filter(magnit < 7) %>% 
  ggplot(aes(x=magnit))+
  geom_histogram(binwidth = 0.1)
```

```{r}
df_filtered %>% 
  ggplot(aes(x=depth,colour=class, y=..density..))+
  geom_freqpoly(binwidth=20)
```


Plot of earthquakes by type

```{r}
ggplot(data = df_filtered) + 
  borders('world', colour='gray50',fill='gray50') +
  geom_point(aes(x=lon, y=lat), alpha = 0.5, size = 1, color = 'blue') +
  facet_wrap(~class)
```

Plot of earthquakes by magnitude
```{r}
ggplot(data = df_filtered) + 
  borders('world', colour='gray50',fill='gray50') +
  geom_point(aes(x=lon, y=lat, color = class), alpha = 0.3, size = 1)+
  scale_colour_manual(values = c("yellow","red", "blue", "black"))
```

Plot magnitude by erruption type
```{r}
ggplot(df_filtered, aes(x = magnit)) +
  geom_density(aes(fill = type), alpha = 0.4)
```


# ```{r}
# ggplot(df_filtered, aes(x = depth, y = magnit)) +
#   geom_point(aes(colour = type, shape = type), alpha = 0.4)
# ```

Plot number of earthquakes per year
```{r}
df_filtered %>% 
  filter(type == 'Earthquake') %>% 
  group_by(year) %>% 
  summarise(events = n()) %>% 
  ggplot(aes(x = year, y = events))+
  geom_bar(stat = 'identity') +
  theme_minimal()
```

Plot average depth based on year
```{r}
df_filtered %>% 
  filter(type == 'Earthquake') %>% 
  group_by(year) %>% 
  summarise(mean_depth = mean(depth)) %>% 
  ggplot(aes(x = year, y = mean_depth))+
  geom_bar(stat = 'identity') +
  theme_minimal()
```

Plot average number of earthquakes for every month
```{r}
df_filtered %>% 
  filter(type == 'Earthquake') %>% 
  group_by(month) %>% 
  summarise(events = n()) %>% 
  ggplot(aes(x = month, y = events))+
  geom_bar(stat = 'identity') +
  theme_minimal()
```

Plot average depth of earthquake based on month
```{r}
df_filtered %>% 
  # filter(type == 'Earthquake') %>% 
  group_by(month) %>% 
  summarise(mean_depth = mean(depth)) %>% 
  ggplot(aes(x = month, y = mean_depth))+
  geom_bar(stat = 'identity') +
  theme_minimal()
```

Animation of earthquakes across years
```{r}
df_filtered %>% 
  plot_ly(
    x = ~lon,
    y = ~lat,
    frame = ~year,
    type = 'scatter',
    mode = 'markers',
    alpha = 0.3
) %>% 
  layout(xaxis = list(range = c(-200, 200)),
         yaxis = list(range = c(-100, 100)))
```

Magnitude against depth
```{r}
df_filtered %>% 
  ggplot(aes(x=depth,y=magnit))+
  geom_point(alpha=0.2)
```

## Modelling

### Polynomial regression (only cubic)
```{r}
cubic_lm <- lm(df_filtered$magnit~df_filtered$depth+I(df_filtered$depth^2)+I(df_filtered$depth^3))
plot(df_filtered$magnit,cubic_lm$fitted.values)
plot(cubic_lm$residuals, cubic_lm$fitted.values)

df_filtered %>% 
  ggplot(aes(x=depth,y=magnit))+
  geom_point(alpha=0.2)
  
cubic <- function(x){
  coef <- cubic_lm$coefficients
  return(coef[1]+x*coef[2]+x**2*coef[3]+x**3*coef[4])
}

df_filtered %>% 
  ggplot(aes(x=depth,y=magnit))+
  geom_point()+
  stat_function(fun = cubic, color='red')
```

Median of pairwise distances rows = observations
```{r}
pairwise_dist_median <- function(x){
  n <- dim(x)[1]
  total_iterations = 0
  dist <- vector(mode = "numeric", (n-1)*(n-2))
  for(i in 1:(n-1)){
    for(j in (i+1):n){
      total_iterations <- total_iterations + 1
      dist[total_iterations] <- norm(x[i,]-x[j,], "2")
    }
  }
  return(list(dist, median(dist)))
}
```

```{r}
rbf <- function(centroid,sigma){
  rbf_in <- function(x) {
    phi <- exp(-(norm(x-centroid, "2")^2/(2 * sigma*sigma)))
    return(phi)
  }
  return(rbf_in)
}
```

We want to use the gaussian rbf as a feature transformation on both data points. 
```{r}
create_data_matrix <- function(X, n) {
  num_cols <- dim(X)[2]
  num_rows <- dim(X)[1]
  #todo fix n = 1
  centroid_indices <- sample(1:num_rows, n)
  centroids <- X[centroid_indices,, drop = FALSE]
  sigma <- 100 # TODO: pairwise distance not working find alternative
  print(sigma)
   
  big_phi_x <- matrix(, nrow=num_rows, ncol=n)
  i <- 0
  for(i in 1:n) {
    centroid <- centroids[i,]
    phi_i <- rbf(centroid, sigma)
    phi_i_x <- apply(X, 1, phi_i)
    big_phi_x[,i] <-  phi_i_x
    i <- i + 1
  }
  return(big_phi_x)
}
```

Parameter estimate
```{r}
W_LS <- function(X, y){
  result <- solve((t(X) %*% X), t(X) %*% y)
  return(result)
}
```



Check Radial Basis
```{r}
num <- 5000
df_filtered_reduced <- df_filtered[1:num, ]
data_matrix <- create_data_matrix(matrix(df_filtered_reduced$depth), 5)
data_matrix <- cbind(rep(1, num), data_matrix)
w_param <- W_LS(data_matrix, df_filtered_reduced$magnit)
```

```{r}

yvals <- data_matrix %*% w_param
xvals <- df_filtered_reduced$depth
mynewdf <- data.frame(x=xvals, y=yvals, z=df_filtered_reduced$magnit)
ggplot(data=mynewdf, aes(x=x,y=y)) +
  geom_line() + 
  geom_point(aes(x=x, y=z), alpha=0.1)

```



```{r}
y_hat <- data_matrix %*% w_param
resid <- (y_hat-df_filtered_reduced$magnit)^2

df_plot <- cbind(df_filtered_reduced$magnit,df_filtered_reduced$depth,resid,y_hat)
colnames(df_plot) <- c("y","x","resid","fitted")

ggplot(df_plot,aes(x,y))
```


